Relating two Abstractions for Formal Array Languages

### 2014-04-11 paper.tex
Use the preprint option to the document class Also remove keywords and classifications
Remove the mention of deforestation
Change fusion example to use same as for MoA

\documentclass{sigplanconf}

@@

\usepackage{amsmath}
\usepackage{longtable}
\usepackage{stmaryrd}
\usepackage[mathletters]{ucs}
\usepackage[utf8x]{inputenc}

@@

\category{CR-number}{subcategory}{third-level}

% general terms are not compulsory anymore, 
% you may leave them out
\terms
term1, term2

\keywords
keyword1, keyword2

@@

\begin{verbatim}
data Pull a = Pull { ixf    :: (Int -> a)
                   , length :: Int
                   }
\end{verbatim}

@@

\begin{figure}
\begin{verbatim}
map :: (a -> b) -> Pull a -> Pull b
map f (Pull ixf l) = Pull (f . ixf) l

@@

enumFromTo :: Int -> Int -> Pull Int
enumFromTo f t = Pull (\i -> i + f) (t - f + 1)
\end{verbatim}
\caption{Example functions on pull arrays}
\label{fig:pullex}
\end{figure}

@@

As mentioned, one of the most important advantages of Pull arrays is
that they support
fusion\cite{gill1993short,axelsson2010feldspar,keller2010regular}. Fusion
is a program transformation which removes intermediate data structures
and is a decendant of deforestation
\cite{wadler1990deforestation}. Fusion supercedes transformations such
as loop fusion in optimizing compilers for imperative languages.  Pull
arrays make it particularly simple to achieve fusion. It simply
amounts to inlining the definitions of functions.

As an example of fusion consider the example
\newline
\verb!map (*2) (map (*5) arr)!, where each element of the array
\verb!arr! is first multiplied by five and then again by
two. Conceptually, there is an intermediate array created after the
first \verb!map! but thanks to fusion it will be removed. Fusion will proceed
as follows:

\begin{verbatim}
map (*2) (map (*5) arr)
  => { arr }
map (*2) (map (*5) (Pull ixf l)
  => { map }
map (*2) (Pull ((*5) . ixf) l)
  => { map }
Pull ((*2) . (*5) . ixf) l)
\end{verbatim}

The result is a single Pull array, the intermediate array has been removed.

>>>>

\documentclass[preprint]{sigplanconf}

@@

\usepackage{amsmath}
\usepackage{longtable}
\usepackage{stmaryrd}
\usepackage[mathletters]{ucs}
\usepackage[utf8x]{inputenc}
\usepackage{fancyvrb}

@@

% \category{CR-number}{subcategory}{third-level}

% general terms are not compulsory anymore, 
% you may leave them out
%\terms
%term1, term2

%\keywords
%Arrays

@@

\begin{small}
\begin{Verbatim}
data Pull a = Pull { ixf    :: (Int -> a)
                   , length :: Int
                   }
\end{Verbatim}
\end{small}

@@

\begin{figure}
\begin{small}
\begin{Verbatim}
map :: (a -> b) -> Pull a -> Pull b
map f (Pull ixf l) = Pull (f . ixf) l

@@

enumFromTo :: Int -> Int -> Pull Int
enumFromTo f t = Pull (\i -> i + f) (t - f + 1)
\end{Verbatim}
\end{small}
\caption{Example functions on pull arrays}
\label{fig:pullex}
\end{figure}

@@

As mentioned, one of the most important advantages of Pull arrays is
that they support
fusion\cite{gill1993short,axelsson2010feldspar,keller2010regular}. Fusion
is a program transformation which removes intermediate data structures.
Fusion supercedes transformations such
as loop fusion in optimizing compilers for imperative languages.  Pull
arrays make it particularly simple to achieve fusion. It simply
amounts to inlining the definitions of functions.

As an example of fusion consider the example from the section which
was used to illustrate $\psi$-reductions. Here is how it is written
using Pull arrays for the one-dimensional case:\\
{\small
\verb!zipWith (*) (take 2 (reverse a)) (drop 1 (reverse a))!}.\\
Conceptually there are many intermediate arrays here, but they will be
removed by fusion. Fusion will proceed by inlining the definition
of the functions as follows:

\begin{small}
\begin{Verbatim}
zipWith (*) (take 2 (reverse a))
            (drop 1 (reverse a))
  => { arr }
zipWith (*) (take 2 (reverse (Pull ixf l))
            (drop 2 (reverse (Pull ixf l)))
  => { reverse * 2 }
zipWith (*) (take 2 (Pull (\i -> ixf (l - i - 1)) l))
            (drop 1 (Pull (\i -> ixf (l - i - 1)) l))
  => { take , drop }
zipWith (*) (Pull (\i -> ixf (l - i - 1)) (min 2 l))
            (Pull (\i -> ixf (l - i)) (max (l - 1) 0))
  => { zipWith }
Pull (\i -> ixf (l - i - 1) * ixf (l - i))
  (min (min 2 l) (max (l - 1) 0))
\end{Verbatim}
\end{small}

The result is a single Pull array, the intermediate array has been
removed. The result is the same in MoA, modulo syntax and the fact
that we here use a one-dimensional array of unknown size.
